\forestclaw: Hybrid forest-of-octrees AMR for hyperbolic conservation laws

### 2013-07-30 docs/parco13/paper.tex
Start with strong scaling for uniform mesh

Grid refinement and coarsening requires interpolation from coarse grids
to newly created fine grids, and the averaging of fine grid data to a newly
created underlying coarse grid.  This operation is rank-local, analogous to the
general dynamic AMR procedures used in \pforest-based FE codes \cite[Fig.\
4]{BursteddeGhattasStadlerEtAl08}.

@@

In both sets of numerical results, we solve a scalar advection
equation,
\begin{equation}
q_t + ({\bf u}\cdot \nabla) q = 0 ,
\end{equation}
where the velocity field ${\bf u} = (u(\xi,\eta,t), v(\xi,\eta,t)$ is
a prescribed function of computational coordinates.  The relevant
numerical parameters that are set in each case include the size of the
patch on each leaf ($m=8$ or $m=16$), and the minimum and maximum
refinement levels, which in turn fix minimum and maximum effective grid
resolutions.

@@

@@

The authors acknowledge valuable discussion with Randy LeVeque, Marsha
Berger, and Hans-Petter Langtangen.  We also acknowledge David
Ketcheson and the KAUST sponsored HPC$^3$ numerics workshop at which
the initial phases of this project were first discussed.  The second
author would like to also acknowledge the Isaac Newton Institute
(Cambridge, UK), where much of the preliminary development work for
\forestclaw was done. The leaf/patch paradigm was independently
presented by B.\ as part of a talk at the SCI Institute, Salt Lake City, Utah,
USA in July 2011.

>>>>

Mesh refinement and coarsening requires interpolation from coarse grids
to newly created fine grids, and the averaging of fine grid data to a newly
created underlying coarse grid.  This operation is rank-local, analogous to the
general dynamic AMR procedures used in \pforest-based FE codes \cite[Fig.\
4]{BursteddeGhattasStadlerEtAl08}.

@@

In both sets of numerical results, we solve a scalar advection
equation,
\begin{equation}
q_t + ({\bf u}\cdot \nabla) q = 0 ,
\end{equation}
where the velocity field ${\bf u} = (u(\xi,\eta,t), v(\xi,\eta,t)$ is
a prescribed function of computational coordinates.  The relevant
numerical parameters that are set in each case include the size of the
patch on each leaf ($m=8$ or $m=16$), and the minimum and maximum
refinement levels, which in turn fix minimum and maximum effective mesh
resolutions.

@@

\begin{figure}
\begin{center}
  \includegraphics[angle=-90,width=.8\columnwidth]{pdf/uni_advance_strong-crop}
  \\[1ex]
  \includegraphics[angle=-90,width=.8\columnwidth]{pdf/uni_exchange_strong-crop}
\end{center}
\caption{Strong scaling of MPI parallelism for advection on a uniform mesh.
  Top: run time of time integration.  Bottom: time spent in local and parallel
  neighbor exchanges.  The number of patches is $2^{2 \ell}$ at level $\ell$.
  The flat lines for the smaller runs are caused by having
  less patches than MPI processes, leaving some of them idle.  For each data point
  we use the maximum time required across all processes.}
\label{fig:uniscale}
\end{figure}%
We begin our analysis of Example 1 with uniform-mesh experiments run on TACC's
\emph{Stampede} supercomputer at different levels of resolution.  First of all,
we use 1 to 16 MPI processes on one 16-core compute node, obtaining a strong
scaling efficiency for the time integration between 97\%--103\% depending on
the resolution (from levels 1 and 8).  Then, we examine processor counts from
16 to 256 using 16 cores per node.  The scaling behavior remains linear for
the time integration itself, and mostly linear for the local and parallel
neighbor exchanges which require about 10\% of the time integration run time;
see Figure~\ref{fig:uniscale}.  Due to space constraints we omit the
corresponding weak scaling results, which are nearly optimal as well.

@@

The authors acknowledge valuable discussion with Randy LeVeque, Marsha
Berger, and Hans-Petter Langtangen.  We also acknowledge David
Ketcheson and the KAUST sponsored HPC$^3$ numerics workshop at which
the initial phases of this project were first discussed.  The second
author would like to also acknowledge the Isaac Newton Institute
(Cambridge, UK), where much of the preliminary development work for
\forestclaw was done. The leaf/patch paradigm was independently
presented by B.\ as part of a talk at the SCI Institute, Salt Lake City, Utah,
USA in July 2011.
\alert{Add TACC Stampede grants}
